---
title: Word Cloud!
author: Martin Gleason, MS
date: '2018-05-10'
slug: word-cloud
categories:
  - word cloud
  - text mining
tags:
  - All Comics Considered
---
```{r, packages, echo = FALSE, message = FALSE}
library("plyr")
library("tidyverse")
library("wordcloud")
library("tidytext")

library("googlesheets")
library("kableExtra")

source("/Users/marty/Dropbox (Personal)/Coding Projects/R/scripts/create_df.R")
source("/Users/marty/Dropbox (Personal)/Coding Projects/R/scripts/answers_from_column.R")

sheets <- gs_ls()

survey_key<- "1WemMnRJphcI5BPgeSAC8wj40If0foT3BH0o-oUApj-g" #need to find a way to automate this right away. 

register_survey <- gs_key(survey_key)
survey_full <- gs_read(register_survey)
survey_anon <- survey_full %>% select(-2)
emails <- survey_full %>% select(2)

#Brew color
spectral <- brewer.pal(8, "Spectral")

#sample text work. a2 is answers to question2. the column was named ansers by names(). the output is worsd
a2 <- survey_anon[2]
names(a2) <- "answers"
```

Today, my [podcast](https://allcomicsconsidered.com) launched our [Patreon](https://www.patreon.com/allcomicsconsidered). I had hoped to finish the write of the 2018 Listener Survey analysis I started, but as it turns out, I had a lot of other projects to write up. 

Most of them probation related.

BUT. In honor of the launch and to get back into the swing of things, I present the word cloud from the ACC 2018 Listener Survey. The image below represent the reasons why people listen to our little show, featuring the `tidytext` package from the amazing [Dr. Julia Silge](https://juliasilge.com/).

```{r, wordcloud, include=TRUE, warning=FALSE, echo=TRUE, message=FALSE}
a2_wordcloud <- a2  %>% unnest_tokens(word, answers) %>% anti_join(stop_words) %>% 
  count(word) %>% filter(n > 2) %>% 
  with(wordcloud(word, n, colors = spectral, random.order = FALSE, random.color = FALSE))
```
